Supplement to `\ourlongtitle'

### 2021-10-13 Final_version/glo-2020-pm-PaRIS.tex
Added final version

\subsection{Pseudo marginalisation in Monte Carlo methods}
\label{sec:pseudo:marginalisation}

@@

Thus, a pointwise estimate of $\ell(x)$ can be obtained by generating $\zeta$ from $\mathbf{R}(x, \rmd z)$ and computing the statistic $\ell \langle \zeta \rangle(x)$, the \emph{pseudo marginal}. In Monte Carlo methods, the operation of replacing, when necessary, the true marginal $\ell$ by its pseudo marginal is referred to as \emph{pseudo marginalisation}. Interestingly, even though pseudo marginalisation is based on the plug-in principle, it preserves typically the consistency of an algorithm. In order to see the this, let $\bar{\mathsf{X}} \eqdef \mathsf{X} \times \mathsf{Z}$ and $\bar{\mathcal{X}} \eqdef \mathcal{X} \tensprod \mathcal{Z}$; then one may define an extended target distribution $\bar{\pi}(\rmd \bar{x}) \eqdef \bar{\lambda}(\rmd \bar{x}) / \bar{\lambda} \1_{\bar{\mathsf{X}}} = \bar{\lambda}(\rmd \bar{x}) / \lambda \1_{\mathsf{X}}$ on $(\bar{\mathsf{X}}, \bar{\mathcal{X}})$, where 

@@

(with $\bar{x} = (x, z)$). By \eqref{eq:pseudo:marginalisation}, $\pi$ is the marginal of $\bar{\pi}$ with respect to the $x$ component. This means that we may produce a random sample $(\xi^i)_{i = 1}^N$ in $\Xset$ targeting $\pi$ by generating a sample $(\xi^i, \zeta^i)_{i = 1}^N$ targeting $\bar{\pi}$ and simply discarding the $\Zset$-valued variables $(\zeta^i)_{i = 1}^N$. Let $\rho$ be a Markov transition density on $\Xset$ with respect to the reference measure $\mu$. Then following \cite{andrieu:robert:2009}, a Markov chain $(\xi_m, \zeta_m)_{m \in \nset}$ targeting $\bar{\pi}$ can be produced on the basis of the Metropolis-Hastings algorithm as follows. Given a state $(\xi_m, \zeta_m)$, a candidate $(\xi^\ast, \zeta^\ast)$ for the next state is generated by drawing $\xi^\ast \sim \rho(x) \, \mu(\rmd x)$ and $\zeta^\ast \sim \kernel{R}(\xi^\ast, \rmd z)$ and accepting the same with probability 

@@

In the present paper we will generalise the pseudo-marginal approach towards biased estimation by allowing the function 
\begin{equation} \label{eq:biased:pseudo:marginalisation}
    \ell^\precpar(x) \eqdef \int \ell \langle z \rangle(x) \, \mathbf{R}(x, \rmd z)
\end{equation}
on $\Xset$ to be different from $\ell$. Here $\varepsilon \geq 0$ is some accuracy parameter describing the distance between $\ell^\precpar$ and $\ell$. Such biased estimates appear naturally, e.g. when the law $\pi$ is governed by a diffusion process and the density $\ell$ is approximated on the basis of different discretisation schemes; see the next section. In that case, $\precpar$ plays the role of the discretisation step size. In \eqref{eq:biased:pseudo:marginalisation}, also the estimator $\ell \langle z \rangle(x)$ and the kernel $\mathbf{R}(x, \rmd z)$ may depend on $\precpar$, even though this is suppressed in the notation. By introducing the possibly unnormalised measure $\lambda^\varepsilon(\rmd x) \eqdef \ell^\precpar(\rmd x) \, \mu(\rmd x)$ on $\Xfd$ we may define the \emph{skew} target probability measure
$$
    \pi^\varepsilon(\rmd x) \eqdef \frac{\lambda^\precpar(\rmd x)}{\lambda^\precpar \1_\Xset}
$$
on $\Xfd$. In the biased case, generating a sample $(\xi^i, \zeta^i, \omega^i)_{i = 1}^N$ targeting \eqref{eq:extended:target} will, by \eqref{eq:biased:pseudo:marginalisation}, provide a sample $(\xi^i, \omega^i)_{i = 1}^N$ targeting $\pi^\varepsilon$ as a by-product. Thus, it is of utmost importance to obtain control over the bias between $\pi$ and $\pi^\precpar$, which is possible under the assumption that there exists a constant $c \geq 0$ such that for all $h \in \bmf{\Xfd}$ and $\varepsilon$,   
\begin{equation} \label{eq:lipschitz:simple:case}
    \left| \lambda^\varepsilon h - \lambda h \right| \leq c \varepsilon \| h \|_\infty. 
\end{equation}
For instance, in the diffusion process case mentioned above, a condition of type \eqref{eq:lipschitz:simple:case} holds, as we will see in Section~\ref{sec:theoretical:results}, typically true in the case where the density is approximated using the \emph{Durham-Gallant estimator} \cite{durham:gallant:2002}. Using that for all $h \in \bmf{\Xfd}$,  
$$
    \pi^\precpar h - \pi h = \pi^\precpar h \left( 1 - \frac{\lambda^\varepsilon \1_\Xset}{\lambda \1_\Xset} \right) + \frac{\lambda^\precpar h - \lambda h}{\lambda \1_\Xset}, 
$$
we straightforwardly obtain the bound  
$$
    \left| \pi^\precpar h - \pi h \right| \leq \precpar \frac{2 c}{\lambda \1_\Xset} \| h \|_\infty
$$
on the systematic error induced by the skew model. Note that the unbiased case \eqref{eq:pseudo:marginalisation} corresponds to letting $\precpar = 0$ in assumption~\eqref{eq:lipschitz:simple:case}. In the next section we will present a solution to the main problem addressed in Section~\ref{sec:model} exploring a pseudo-marginalised version of the PaRIS discussed in Section~\ref{sec:PaRIS}.

@@

%Let us turn our focus to backward sampling. 
As intractability of $\ud{n}$ implies intractability of the kernel $\trm{n}$ (defined in \eqref{eq:def:trm}), we aim at further pseudo marginalisation by embedding $\trm{n}$ into the extended probability kernel  
$$
\trmext{n}(i, j, \rmd z) \propto \ewght{n}{j} \ukest{n}{z}(\epart{n}{j}, \epart{n + 1}{i}) \, \ukdist{n}(\epart{n}{j}, \epart{n + 1}{i}, \rmd z)
$$
on $\intvect{1}{\N} \times \powerset{\intvect{1}{\N}} \tensprod \Zfd_{n + 1}$. For every $i$, the marginal of $\trmext{n}(i, \cdot)$ with respect to the $j$ component is, by \eqref{eq:def:udmod}, proportional to $\ewght{n}{j} \udmod{n}(\epart{n}{j}, \epart{n + 1}{i})$, a distribution that we expect to be close to $\trm{n}(i, \cdot)$ for small $\precpar$. The intractable sampling step on Line~3 in Algorithm~\ref{alg:ideal:BS} can therefore be replaced by sampling from $\trmext{n}(i, \cdot)$, after which the auxiliary variables are discarded. The latter sampling operation will be examined in detail in the next section. This approach, which we express concisely as 
$$
    (\tstat[i]{n + 1})_{i = 1}^\N \sim \mathsf{pmBS}((\epart{n}{i}, \tstat[i]{n}, \ewght{n}{i})_{i = 1}^\N, (\epart{n + 1}{i})_{i = 1}^\N), 
$$
is summarised in Algorithm~\ref{alg:pm:backward:sampling}.

@@

Assume that there exists some measurable nonnegative function $c$ on $\Xset_{n + 1}$ such that for all $(x_{n:n + 1}, z) \in \Xset_n \times \Xset_{n + 1} \times \Zset_{n + 1}$,
$
\ukest{n}{z}(x_n, x_{n + 1}) \leq c(x_{n + 1}).  
$
Since this condition allows the Radon--Nikodym derivative of $\trmext{n}(i, \cdot)$ with respect to the probability measure 
\begin{equation} \label{eq:proposal:pm:rejection}
\rho_n^i(j, \rmd z) \propto \ewght{n}{j} \kernel{R}(\epart{n}{j}, \epart{n + 1}{i}, \rmd z)
\end{equation} 
on $\powerset{\intvect{1}{\N}} \tensprod \Zfd_{n + 1}$ to be bounded uniformly as 
$$
\frac{\rmd \trmext{n}(i, \cdot)}{\rmd \rho_n^i}(j, z) \leq \frac{c(\epart{n + 1}{i}) \sumwght{n}}{\sum_{i' = 1}^\N \ewght{n}{i'} \udmod{n}(\epart{n}{i'}, \epart{n + 1}{i})},  
$$ 
we may sample from the target $\trmext{n}(i, \cdot)$ using rejection sampling. Thus, the following procedure is iterated until acceptance: simulate a candidate $(J^\ast, \zeta^\ast)$ from $\rho_n^i$ by drawing $J^\ast \sim \cat(\{ \ewght{n}{j} \}_{j = 1}^\N)$ and $\zeta^\ast \sim \kernel{R}(\epart{n}{J^\ast}, \epart{n + 1}{i}, \rmd z)$; then accept the same with (tractable) probability 
$$
\accprobext \eqdef \frac{\ukest{n}{\zeta^\ast}(\epart{n}{J^\ast}, \epart{n + 1}{i})}{c(\epart{n + 1}{i})}. 
$$ 
Then conditionally to acceptance, the candidate has distribution $\trmext{n}(i, \cdot)$.  
Notably, the probability $\accprobext$ is obtained by simply plugging a transition density estimate provided by \hypref{assum:biased:estimate} into the probability $\accprob$ (see \eqref{eq:std:acc:prob:backward:sampling}) corresponding to the case where $\ud{n}$ is known. Moreover, since the proposal density is independent of $i$, the expected complexity of this sampling schedule is linear in the number of particles (we refer again to \cite{douc:garivier:moulines:olsson:2010}).

@@

In some cases, bounding the estimator $\ukest{n}{z}(x_n, x_{n + 1})$ uniformly in $z$ and $x_n$ is not possible. Still, we may sample from $\trmext{n}(i, \cdot)$ using the Metropolis-Hastings algorithm with $\rho_n$ (in \eqref{eq:proposal:pm:rejection}) as independent proposal. In this case, $(\bi{n + 1}{i}{j}, \zpart{n + 1}{(i, j)})_{j = 1}^{\K}$ is a Markov chain generated recursively by the following mechanism. Given a state $\bi{n + 1}{i}{j} = J$ and $\zpart{n + 1}{(i, j)} = \zeta$, a candidate $(J^\ast, \zeta^\ast)$ for the next state is drawn from $\rho_n$ (as described above) and accepted with probability 
$$
\accprobext[MH] \eqdef 1 \wedge \frac{\ukest{n}{\zeta^\ast}(\epart{n}{J^\ast}, \epart{n + 1}{i})}{\ukest{n}{\zeta}(\epart{n}{J}, \epart{n + 1}{i})}. 
$$  
In the case of rejection, the next state is assigned the previous state. The resulting Markov chain has $\trmext{n}(i, \cdot)$ as stationary distribution and similar to the case of rejection sampling, the acceptance probability $\accprobext[MH]$ can be viewed as a plug-in estimate of the corresponding probability $\accprob[MH]$ (see \eqref{eq:std:MH:prob:backward:sampling}).

@@

\begin{example}[Durham--Gallant estimator]
\label{eq:durham:gallant}
As an illustration, we return to the state-space model framework discussed in Example~\ref{ex:state-space:models}. Let $\Xset \eqdef \rset^{d_x}$ and $\Yset \eqdef \rset^{d_y}$ be equipped with their respective Borel $\sigma$-fields $\Xfd$ and $\Yfd$, and let $(X_t)_{t > 0}$ be some diffusion process on $\Xset$ driven by the homogeneous stochastic differential equation
\begin{equation} \label{eq:SDE}
\rmd X_t = \mu(X_t) \, \rmd t + \sigma(X_t) \, \rmd W_t, \quad t > 0, 
\end{equation}
where $X_0 = 0$, $(W_t)_{t > 0}$ is $d_x$-dimensional Brownian motion, $b : \Xset \to \Xset$ and $\sigma : \Xset \to \rset^{d_x \times d_x}$ are twice differentiable with bounded first and second order derivatives. In addition, the matrix $\sigma \sigma^\intercal$ is uniformly non-degenerate. Let $(\mathcal{F}_t)_{t > 0}$ be the natural filtration generated by the process $(X_t)_{t > 0}$. The state sequence $(X_t)_{t > 0}$ is latent but partially observed at discrete time points $(t_n)_{n \in \nsetpos}$ which are assumed to be equally spaced for simplicity, \ie, $t_n = t_1 + \delta (n - 1)$ for all $n$ and some $\delta > 0$. Abusing notations, we denote $X_n \eqdef X_{t_n}$ and let $q_\delta$ be the transition density of $(X_n)_{n \in \nsetpos}$. Denote by $\kernel{Q}_\delta$ the transition kernel induced by $q_\delta$. In general, $q_\delta$ is intractable, which makes the problem of computing online, for a given data stream $(y_n)_{n \in \nsetpos}$ in $\Yset$, the sequence of joint-smoothing distributions \eqref{eq:smooth} in models of this sort very challenging. Still, using the Euler scheme, one may, for small $\delta$, approximate $q_\delta$ by 
$$
\bar{q}_\delta(x_n, x_{n + 1}) \eqdef \phi(x_{n + 1}; x_n + \delta \mu(x_n), \delta \sigma^2(x_n)), 
$$ 
where $\phi(\cdot; m, s^2)$ is the density of the Gaussian distribution with mean $m$ and variance $s^2$. Let $\bar{\kernel{Q}}_\delta$ be the transition kernel induced by $\bar{q}_\delta$. Since the approximation $\bar{q}_\delta$ is poor for $\delta$ not small enough, we may instead, as suggested in \cite{durham:gallant:2002}, pick some finer step size $\precpar \in \precparsp_\delta \eqdef \{ \delta / n : n \in \nsetpos \}$ and estimate the density $q_\delta(x_n, x_{n + 1})$ by $q_\delta \langle \zeta \rangle (x_n, x_{n + 1})$, where $\zeta = (\zeta^i)_{i = 1}^L$ are independent draws from some proposal $r(x_n, x_{n + 1}, z) \, \rmd z$ on $\Xset^2 \times \Xfd^{\delta / \precpar - 1}$ and  
$$
q_\delta \langle z \rangle (x_n, x_{n + 1}) \eqdef \frac{1}{L} \sum_{i = 1}^L \frac{\prod_{k = 1}^{\delta / \precpar} \bar{q}_\precpar(z_{k - 1}^i, z_k^i)}{r(x_n, x_{n + 1}, z^i)},  
$$
with $z^i = (z_1^i, \ldots, z_{\delta / \precpar - 1}^i)$ and, by convention, $z_0^i = x_n$ and $z_{\delta / \precpar}^i = x_{n + 1}$. In \cite{durham:gallant:2002}, $r(x_n, x_{n + 1}, z) \, \rmd z$ is the distribution of a discretised (possibly modified) \emph{Brownian bridge}, \ie, Brownian motion started at $x_n$ and conditioned to terminate at $x_{n + 1}$.

@@

\begin{example} 
[approximate Bayesian computation smoothing]
Consider joint smoothing in a fully dominated general state-space HMM for which the state likelihood functions $(\md{n}(\cdot, y_n))_{n \in \nset}$ are intractable (or expensive to evaluate) for any given sequence $(y_n)_{n \in \nset}$ of observations in $\rset^{d_y}$. In the case where it is possible (or faster) to sample observation emissions according to the kernels $(\mk_n)_{n \in \nset}$ one may then 
take an \emph{approximate-Bayesian-computation} (ABC) approach (see {\eg}  \cite{marin:pudlo:robert:ryder:2012}), and replace any value $\md{n}(x_n, y_n)$ by a point estimate $\kappa_\precpar(\zeta_n - y_n)$, where $\zeta_n \sim \mk_n(x_n, \cdot)$ and $\kappa_\varepsilon$ is a $d_y$-dimensional kernel density scaled by some bandwidth $\precpar > 0$. In \cite{martin:jasra:singh:whiteley:delmoral:maccoy:2014}, the authors apply the forward-only smoothing approach of \cite{delmoral:doucet:singh:2010} to this approximate model, yielding a particle-based ABC smoothing algorithm. Also this framework is covered by \hypref{assum:biased:estimate}, by letting $\ukest{n}{z}(x_n, x_{n + 1}) = \hd{n}(x_n, x_{n + 1}) \kappa_\precpar(z - y_{n + 1})$ and $\ukdist{n}(x_n, x_{n + 1}, \rmd z) = \mk_n(x_{n + 1}, \rmd z)$. In this case, 
\begin{equation} \label{eq:ukmod:ABC:smoothing}
\ukmod{n}(x_n, \rmd x_{n + 1}) = \hk_n(x_n, \rmd x_{n + 1}) \int \kappa_\precpar(z - y_{n + 1}) \, \mk_n(x_{n + 1}, \rmd z). 
\end{equation}
\end{example}

>>>>

\subsection{Pseudo-marginalisation in Monte Carlo methods}
\label{sec:pseudo:marginalisation}

@@

Thus, a pointwise estimate of $\ell(x)$ can be obtained by generating $\zeta$ from $\mathbf{R}(x, \rmd z)$ and computing the statistic $\ell \langle \zeta \rangle(x)$, the \emph{pseudo-marginal}. In Monte Carlo methods, the operation of replacing, when necessary, the true marginal $\ell$ by its pseudo-marginal is referred to as \emph{pseudo-marginalisation}. Interestingly, even though pseudo-marginalisation is based on the plug-in principle, it preserves typically the consistency of an algorithm. In order to see the this, let $\bar{\mathsf{X}} \eqdef \mathsf{X} \times \mathsf{Z}$ and $\bar{\mathcal{X}} \eqdef \mathcal{X} \tensprod \mathcal{Z}$; then one may define an extended target distribution $\bar{\pi}(\rmd \bar{x}) \eqdef \bar{\lambda}(\rmd \bar{x}) / \bar{\lambda} \1_{\bar{\mathsf{X}}} = \bar{\lambda}(\rmd \bar{x}) / \lambda \1_{\mathsf{X}}$ on $(\bar{\mathsf{X}}, \bar{\mathcal{X}})$, where 

@@

(with $\bar{x} = (x, z)$). By \eqref{eq:pseudo:marginalisation}, $\pi$ is the marginal of $\bar{\pi}$ with respect to the $x$ component. This means that we may produce a random sample $(\xi^i)_{i = 1}^N$ in $\Xset$ targeting $\pi$ by generating a sample $(\xi^i, \zeta^i)_{i = 1}^N$ targeting $\bar{\pi}$ and simply discarding the $\Zset$-valued variables $(\zeta^i)_{i = 1}^N$. Let $\rho$ be a Markov transition density on $\Xset$ with respect to the reference measure $\mu$. Then following \cite{andrieu:robert:2009}, a Markov chain $(\xi_m, \zeta_m)_{m \in \nset}$ targeting $\bar{\pi}$ can be produced on the basis of the Metropolis-Hastings algorithm as follows. Given a state $(\xi_m, \zeta_m)$, a candidate $(\xi^\ast, \zeta^\ast)$ for the next state is generated by drawing $\xi^\ast \sim \rho(\xi_m, x) \, \mu(\rmd x)$ and $\zeta^\ast \sim \kernel{R}(\xi^\ast, \rmd z)$ and accepting the same with probability 

@@

In the present paper we will generalise the pseudo-marginal approach towards biased estimation by allowing the function 
\begin{equation} \label{eq:biased:pseudo:marginalisation}
    \ell^\precpar(x) \eqdef \int \ell \langle z \rangle(x) \, \mathbf{R}(x, \rmd z)
\end{equation}
on $\Xset$ to be different from $\ell$. Here $\varepsilon \geq 0$ is some accuracy parameter describing the distance between $\ell^\precpar$ and $\ell$. Such biased estimates appear naturally, \eg\ when the law $\pi$ is governed by a diffusion process and the density $\ell$ is approximated on the basis of different discretisation schemes; see the next section. In that case, $\precpar$ plays the role of the discretisation step size. In \eqref{eq:biased:pseudo:marginalisation}, also the estimator $\ell \langle z \rangle(x)$ and the kernel $\mathbf{R}(x, \rmd z)$ may depend on $\precpar$, even though this is suppressed in the notation. By introducing the possibly unnormalised measure $\lambda^\varepsilon(\rmd x) \eqdef \ell^\precpar(\rmd x) \, \mu(\rmd x)$ on $\Xfd$ we may define the \emph{skew} target probability measure
$$
    \pi^\varepsilon(\rmd x) \eqdef \frac{\lambda^\precpar(\rmd x)}{\lambda^\precpar \1_\Xset}
$$
on $\Xfd$. In the biased case, generating a sample $(\xi^i, \zeta^i, \omega^i)_{i = 1}^N$ targeting \eqref{eq:extended:target} will, by \eqref{eq:biased:pseudo:marginalisation}, provide a sample $(\xi^i, \omega^i)_{i = 1}^N$ targeting $\pi^\varepsilon$ as a by-product. Thus, it is of utmost importance to obtain control over the bias between $\pi$ and $\pi^\precpar$, which is possible under the assumption that there exists a constant $c \geq 0$ such that for all $h \in \bmf{\Xfd}$ and $\varepsilon$,   
\begin{equation} \label{eq:lipschitz:simple:case}
    \left| \lambda^\varepsilon h - \lambda h \right| \leq c \varepsilon \| h \|_\infty. 
\end{equation}
For instance, in the diffusion process case mentioned above, a condition of type \eqref{eq:lipschitz:simple:case} holds, as we will see in Section~\ref{sec:theoretical:results}, typically true in the case where the density is approximated using the \emph{Durham-Gallant estimator} \cite{durham:gallant:2002}. Using that for all $h \in \bmf{\Xfd}$,  
$$
    \pi^\precpar h - \pi h = \pi^\precpar h \left( 1 - \frac{\lambda^\varepsilon \1_\Xset}{\lambda \1_\Xset} \right) + \frac{\lambda^\precpar h - \lambda h}{\lambda \1_\Xset}, 
$$
we straightforwardly obtain, using \eqref{eq:lipschitz:simple:case}, the bound  
$$
    \left| \pi^\precpar h - \pi h \right| \leq \precpar \frac{2 c}{\lambda \1_\Xset} \| h \|_\infty
$$
on the systematic error induced by the skew model. Note that the unbiased case \eqref{eq:pseudo:marginalisation} corresponds to letting $\precpar = 0$ in assumption~\eqref{eq:lipschitz:simple:case}. In the next section we will present a solution to the main problem addressed in Section~\ref{sec:model} exploring a pseudo-marginalised version of the PaRIS discussed in Section~\ref{sec:PaRIS}.

@@

%Let us turn our focus to backward sampling. 
As intractability of $\ud{n}$ implies intractability of the kernel $\trm{n}$ (defined in \eqref{eq:def:trm}), we aim at further pseudo-marginalisation by embedding $\trm{n}$ into the extended Markov kernel  
$$
\trmext{n}(i, j, \rmd z) \propto \ewght{n}{j} \ukest{n}{z}(\epart{n}{j}, \epart{n + 1}{i}) \, \ukdist{n}(\epart{n}{j}, \epart{n + 1}{i}, \rmd z)
$$
on $\intvect{1}{\N} \times \powerset{\intvect{1}{\N}} \tensprod \Zfd_{n + 1}$. For every $i$, the marginal of $\trmext{n}(i, \cdot)$ with respect to the $j$ component is, by \eqref{eq:def:udmod}, proportional to $\ewght{n}{j} \udmod{n}(\epart{n}{j}, \epart{n + 1}{i})$, a distribution that we expect to be close to $\trm{n}(i, \cdot)$ for small $\precpar$. The intractable sampling step on Line~3 in Algorithm~\ref{alg:ideal:BS} can therefore be replaced by sampling from $\trmext{n}(i, \cdot)$, after which the auxiliary variables are discarded. The latter sampling operation will be examined in detail in the next section. This approach, which we express concisely as 
$$
    (\tstat[i]{n + 1})_{i = 1}^\N \sim \mathsf{pmBS}((\epart{n}{i}, \tstat[i]{n}, \ewght{n}{i})_{i = 1}^\N, (\epart{n + 1}{i})_{i = 1}^\N), 
$$
is summarised in Algorithm~\ref{alg:pm:backward:sampling}.

@@

Assume that there exists some measurable nonnegative function $c$ on $\Xset_{n + 1}$ such that for all $(x_{n:n + 1}, z) \in \Xset_n \times \Xset_{n + 1} \times \Zset_{n + 1}$,
$
\ukest{n}{z}(x_n, x_{n + 1}) \leq c(x_{n + 1}).  
$
Since this condition allows the Radon--Nikodym derivative of $\trmext{n}(i, \cdot)$ with respect to the probability measure 
\begin{equation} \label{eq:proposal:pm:rejection}
\rho_n^i(j, \rmd z) \propto \ewght{n}{j} \kernel{R}(\epart{n}{j}, \epart{n + 1}{i}, \rmd z)
\end{equation} 
on $\powerset{\intvect{1}{\N}} \tensprod \Zfd_{n + 1}$ to be bounded uniformly as 
$$
\frac{\rmd \trmext{n}(i, \cdot)}{\rmd \rho_n^i}(j, z) \leq \frac{c(\epart{n + 1}{i}) \sumwght{n}}{\sum_{i' = 1}^\N \ewght{n}{i'} \udmod{n}(\epart{n}{i'}, \epart{n + 1}{i})},  
$$ 
we may sample from the target $\trmext{n}(i, \cdot)$ using rejection sampling. Thus, the following procedure is iterated until acceptance: simulate a candidate $(J^\ast, \zeta^\ast)$ from $\rho_n^i$ by drawing $J^\ast \sim \cat(\{ \ewght{n}{j} \}_{j = 1}^\N)$ and $\zeta^\ast \sim \kernel{R}(\epart{n}{J^\ast}, \epart{n + 1}{i}, \rmd z)$; then accept the same with (tractable) probability 
$$
\accprobext \eqdef \frac{\ukest{n}{\zeta^\ast}(\epart{n}{J^\ast}, \epart{n + 1}{i})}{c(\epart{n + 1}{i})}. 
$$ 
Then conditionally on acceptance, the candidate has distribution $\trmext{n}(i, \cdot)$.  
Notably, the probability $\accprobext$ is obtained by simply plugging a transition-density estimate provided by \hypref{assum:biased:estimate} into the probability $\accprob$ (see \eqref{eq:std:acc:prob:backward:sampling}) corresponding to the case where $\ud{n}$ is known. Moreover, since the proposal density is independent of $i$, the expected complexity of this sampling schedule is linear in the number of particles (we refer again to \cite{douc:garivier:moulines:olsson:2010}).

@@

In some cases, bounding the estimator $\ukest{n}{z}(x_n, x_{n + 1})$ uniformly in $z$ and $x_n$ is not possible. Still, we may sample from $\trmext{n}(i, \cdot)$ using the Metropolis--Hastings algorithm with $\rho_n$ (in \eqref{eq:proposal:pm:rejection}) as independent proposal. In this case, $(\bi{n + 1}{i}{j}, \zpart{n + 1}{(i, j)})_{j = 1}^{\K}$ is a Markov chain generated recursively by the following mechanism. Given a state $\bi{n + 1}{i}{j} = J$ and $\zpart{n + 1}{(i, j)} = \zeta$, a candidate $(J^\ast, \zeta^\ast)$ for the next state is drawn from $\rho_n$ (as described above) and accepted with probability 
$$
\accprobext[MH] \eqdef 1 \wedge \frac{\ukest{n}{\zeta^\ast}(\epart{n}{J^\ast}, \epart{n + 1}{i})}{\ukest{n}{\zeta}(\epart{n}{J}, \epart{n + 1}{i})}. 
$$  
In the case of rejection, the next state is assigned the previous state. The resulting Markov chain has $\trmext{n}(i, \cdot)$ as stationary distribution and similar to the case of rejection sampling, the acceptance probability $\accprobext[MH]$ can be viewed as a plug-in estimate of the corresponding probability $\accprob[MH]$ (see \eqref{eq:std:MH:prob:backward:sampling}).

@@

\begin{example}[Durham--Gallant estimator]
\label{eq:durham:gallant}
As an illustration, we return to the state-space model framework discussed in Example~\ref{ex:state-space:models}. Let $\Xset \eqdef \rset^{d_x}$ and $\Yset \eqdef \rset^{d_y}$ be equipped with their respective Borel $\sigma$-fields $\Xfd$ and $\Yfd$, and let $(X_t)_{t > 0}$ be some diffusion process on $\Xset$ driven by the homogeneous stochastic differential equation
\begin{equation} \label{eq:SDE}
\rmd X_t = \mu(X_t) \, \rmd t + \sigma(X_t) \, \rmd W_t, \quad t > 0, 
\end{equation}
where $X_0 = 0$, $(W_t)_{t > 0}$ is $d_x$-dimensional Brownian motion, and $b : \Xset \to \Xset$ and $\sigma : \Xset \to \rset^{d_x \times d_x}$ are twice differentiable with bounded first and second order derivatives. In addition, the matrix $\sigma \sigma^\intercal$ is uniformly non-degenerate. Let $(\mathcal{F}_t)_{t > 0}$ be the natural filtration generated by the process $(X_t)_{t > 0}$. The state sequence $(X_t)_{t > 0}$ is latent but partially observed at discrete time points $(t_n)_{n \in \nsetpos}$ which are assumed to be equally spaced for simplicity, \ie, $t_n = t_1 + \delta (n - 1)$ for all $n$ and some $\delta > 0$. Abusing notations, we denote $X_n \eqdef X_{t_n}$ and let $q_\delta$ be the transition density of $(X_n)_{n \in \nsetpos}$. Denote by $\kernel{Q}_\delta$ the transition kernel induced by $q_\delta$. In general, $q_\delta$ is intractable, which makes the problem of computing online, for a given data stream $(y_n)_{n \in \nsetpos}$ in $\Yset$, the sequence of joint-smoothing distributions \eqref{eq:smooth} in models of this sort very challenging. Still, using the Euler scheme, one may, for small $\delta$, approximate $q_\delta$ by 
$$
\bar{q}_\delta(x_n, x_{n + 1}) \eqdef \phi(x_{n + 1}; x_n + \delta \mu(x_n), \delta \sigma^2(x_n)), 
$$ 
where $\phi(\cdot; m, s^2)$ is the density of the Gaussian distribution with mean $m$ and variance $s^2$. Let $\bar{\kernel{Q}}_\delta$ be the transition kernel induced by $\bar{q}_\delta$. Since the approximation $\bar{q}_\delta$ is poor for $\delta$ not small enough, we may instead, as suggested in \cite{durham:gallant:2002}, pick some finer step size $\precpar \in \precparsp_\delta \eqdef \{ \delta / n : n \in \nsetpos \}$ and estimate the density $q_\delta(x_n, x_{n + 1})$ by $q_\delta \langle \zeta \rangle (x_n, x_{n + 1})$, where $\zeta = (\zeta^i)_{i = 1}^L$ are independent draws from some proposal $r(x_n, x_{n + 1}, z) \, \rmd z$ on $\Xset^2 \times \Xfd^{\delta / \precpar - 1}$ and  
$$
q_\delta \langle z \rangle (x_n, x_{n + 1}) \eqdef \frac{1}{L} \sum_{i = 1}^L \frac{\prod_{k = 1}^{\delta / \precpar} \bar{q}_\precpar(z_{k - 1}^i, z_k^i)}{r(x_n, x_{n + 1}, z^i)},  
$$
with $z^i = (z_1^i, \ldots, z_{\delta / \precpar - 1}^i)$ and, by convention, $z_0^i = x_n$ and $z_{\delta / \precpar}^i = x_{n + 1}$. In \cite{durham:gallant:2002}, $r(x_n, x_{n + 1}, z) \, \rmd z$ is the distribution of a discretised (possibly modified) \emph{Brownian bridge}, \ie, Brownian motion started at $x_n$ and conditioned to terminate at $x_{n + 1}$.

@@

\begin{example} 
[approximate Bayesian computation smoothing]
Consider joint smoothing in a fully dominated general state-space HMM for which the state likelihood functions $(\md{n}(\cdot, y_n))_{n \in \nset}$ are intractable (or expensive to evaluate) for any given sequence $(y_n)_{n \in \nset}$ of observations in $\rset^{d_y}$. In the case where it is possible (or faster) to sample observation emissions according to the kernels $(\mk_n)_{n \in \nset}$ one may then 
take an \emph{approximate Bayesian computation} (ABC) approach (see {\eg}  \cite{marin:pudlo:robert:ryder:2012}), and replace any value $\md{n}(x_n, y_n)$ by a point estimate $\kappa_\precpar(\zeta_n - y_n)$, where $\zeta_n \sim \mk_n(x_n, \cdot)$ and $\kappa_\varepsilon$ is a $d_y$-dimensional kernel density scaled by some bandwidth $\precpar > 0$. In \cite{martin:jasra:singh:whiteley:delmoral:maccoy:2014}, the authors apply the forward-only smoothing approach of \cite{delmoral:doucet:singh:2010} to this approximate model, yielding a particle-based ABC smoothing algorithm. Also this framework is covered by \hypref{assum:biased:estimate} by letting $\ukest{n}{z}(x_n, x_{n + 1}) = \hd{n}(x_n, x_{n + 1}) \kappa_\precpar(z - y_{n + 1})$ and $\ukdist{n}(x_n, x_{n + 1}, \rmd z) = \mk_n(x_{n + 1}, \rmd z)$. In this case, 
\begin{equation} \label{eq:ukmod:ABC:smoothing}
\ukmod{n}(x_n, \rmd x_{n + 1}) = \hk_n(x_n, \rmd x_{n + 1}) \int \kappa_\precpar(z - y_{n + 1}) \, \mk_n(x_{n + 1}, \rmd z). 
\end{equation}
\end{example}
